### 2013-05-19 draft.tex
ready to send to Dave and Seth

Our clumped isotope measurements indicate that Silurian tropical oceans were similar to modern tropical oceans in their temperature and \deltao O. Modern measurements of sea surface temperature in the Solomon Sea, east of Papua New Guinea, are 30\degrees C \citep{Reynolds1994}, matching our estimate of the temperature at a similar latitude during the Silurian. The average \deltao O of the modern ocean is approximately $-0.3\permil$ \citep{Shackleton1974}, but without ice, the average \deltao O would decrease to $-1.4\permil$ \citep{Lhomme2005}. The volume of continental glaciers is poorly constrained for the Early Silurian, but regardless, the \deltao O of Silurian oceans was likely within a few permil of the modern value.

The results of this study are also consistent with those of earlier clumped isotope studies. \cite{Came2007} and \cite{Finnegan2011} both measured the clumped isotope compositions of brachiopods from the Early Silurian on Anticosti Island, Canada. These two studies report temperatures of $30-35$\degrees C and seawater \deltao O$_{\text{SMOW}}$ values of around $-1\permil$, which are similar to the results of this study. Importantly, Anticosti Island recorded deposition on the eastern margin of the Panthalassic Ocean during the Silurian, while Gotland recorded deposition in the Rheic Ocean, which was separated from the Panthalassic Ocean by a mountain range. The strong similarity in temperatures and \deltao O between the two studies indicates that both studies likely reflect water that is representative of the global ocean, rather than an isolated inland sea.

Our results, along with those of previous clumped isotope studies \citep{Came2007,Dennis2013,Finnegan2011}, provide consistent explanations for the low \deltao O values of early Paleozoic carbonates. Carbonate clumped isotopes constrain the contribution of a change in seawater \deltao O to only a few permil of the observed trend, ruling out a value of $-8\permil$ for the \deltao O of early Paleozoic oceans \citep{Grossman2012}. Our results also serve as a reminder that many of the fossils preserved in the geologic record lived in tropical oceans. The \deltao O of carbonate in equilibrium with tropical ocean water should not be compared to the \deltao O of carbonate in equilibrium with water of average modern sea surface temperature, because tropical oceans are much warmer than the average sea surface temperature. Clumped isotopes additionally demonstrate that extremely hot ($\sim 70$\degrees C) ocean temperatures cannot be the cause of the depleted \deltao O values of early Paleozoic carbonates. Finally, clumped isotopes, as a highly sensitive indicator of diagenetic alteration, demonstrate that even among samples with original calcite fabric and pristine trace metal signatures, there is a large range in preservation quality. This means that the average fossil calcite in almost any sample suite is likely altered from its primary composition, and that diagenetic alteration is a key contributor to the low \deltao O values of early Paleozoic carbonates.

@@

The results from this study join those of previous clumped isotope studies in implying that the \deltao O of seawater has been fairly constant throughout the Phanerozoic, at around $-1\permil$. Several studies have suggested that high-temperature alteration of the seafloor acts as a buffer for the \deltao O of seawater \citep{Muehlenbachs1976,Gregory1981,Gregory1991,Muehlenbachs1998}, but a more recent model implied that this buffer may not be strong enough to overcome slow changes in the \deltao O of seawater \citep{Jaffres2007}. If this buffer cannot explain the relative constancy of seawater \deltao O values through time, then the two competing controls on seawater \deltao O, alteration of rocks at high and low temperatures, must have remained in balance over the last 500 million years. This suggests that the magnitudes of high-temperature hydrothermal alteration and low-temperature weathering have not changed substantially, or that any changes in the magnitude of one process are tied through an as-yet unidentified feedback loop to balancing changes in the other.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{Fig_veizer-01.jpg}
\caption{Caption}
\label{veizer}
\end{figure}

>>>>

Our clumped isotope measurements indicate that Silurian tropical oceans were similar to modern tropical oceans in their temperature and \deltao O. Modern measurements of sea surface temperature in the Solomon Sea, east of Papua New Guinea, are 30\degrees C \citep{Reynolds1994}, matching our estimate of the temperature at a similar latitude during the Silurian. The average \deltao O$_{\text{SMOW}}$ of the modern ocean is approximately $-0.3\permil$ \citep{Shackleton1974}, but without ice, the average \deltao O$_{\text{SMOW}}$ would decrease to $-1.4\permil$ \citep{Lhomme2005}. The volume of continental glaciers is poorly constrained for the Early Silurian, but regardless, the \deltao O of Silurian oceans was likely within a few permil of the modern value.

The results of this study are also consistent with those of earlier clumped isotope studies. \cite{Came2007} and \cite{Finnegan2011} both measured the clumped isotope compositions of brachiopods from the Early Silurian on Anticosti Island, Canada. These two studies report temperatures of 30 to 35\degrees C and seawater \deltao O$_{\text{SMOW}}$ values of around $-1\permil$, which are similar to the results of this study. Importantly, Anticosti Island recorded deposition on the eastern margin of the Panthalassic Ocean during the Silurian, while Gotland recorded deposition in the Rheic Ocean, which was separated from the Panthalassic Ocean by a mountain range. The strong similarity in temperatures and \deltao O between the two studies indicates that both studies likely reflect water that is representative of the global ocean, rather than an isolated inland sea.

Our results, along with those of previous clumped isotope studies \citep{Came2007,Dennis2013,Finnegan2011}, provide consistent explanations for the low \deltao O values of early Paleozoic carbonates. Carbonate clumped isotopes constrain the contribution of a change in seawater \deltao O to only a few permil of the observed trend, ruling out a value of $-8\permil$ for the \deltao O$_{\text{SMOW}}$ of early Paleozoic oceans \citep{Grossman2012}. Our results also serve as a reminder that many of the fossils preserved in the geologic record lived in tropical oceans. The \deltao O of carbonate in equilibrium with tropical ocean water should not be compared to the \deltao O of carbonate in equilibrium with water of average modern sea surface temperature, because tropical oceans are much warmer than the average sea surface temperature. Clumped isotopes additionally demonstrate that extremely hot ($>50$\degrees C) ocean temperatures cannot be the cause of the depleted \deltao O values of early Paleozoic carbonates. Finally, clumped isotopes, as a highly sensitive indicator of diagenetic alteration, demonstrate that even among samples with original calcite fabric and pristine trace metal signatures, there is a large range in preservation quality. This means that the average fossil calcite in almost any sample suite is likely altered from its primary composition, and that diagenetic alteration is a key contributor to the low \deltao O values of early Paleozoic carbonates.

@@

The results from this study join those of previous clumped isotope studies in implying that the \deltao O$_{\text{SMOW}}$ of seawater has been fairly constant throughout the Phanerozoic, at around $-1\permil$. Several studies have suggested that high-temperature alteration of the seafloor acts as a buffer for the \deltao O of seawater \citep{Muehlenbachs1976,Gregory1981,Gregory1991,Muehlenbachs1998}, but a more recent model implied that this buffer may not be strong enough to overcome slow changes in the \deltao O of seawater \citep{Jaffres2007}. If this buffer cannot explain the relative constancy of seawater \deltao O values through time, then the two competing controls on seawater \deltao O, alteration of rocks at high and low temperatures, must have remained in balance over the last 500 million years. This suggests that the magnitudes of high-temperature hydrothermal alteration and low-temperature weathering have not changed substantially, or that any changes in the magnitude of one process are tied through an as-yet unidentified feedback loop to balancing changes in the other.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{Fig_veizer-01.jpg}
\caption{In gray, the \deltao O$_{\text{PDB}}$ of apparently well-preserved skeletal calcites, compiled by \cite{Veizer1999}. In black are two different hypotheses for the trend in the \deltao O$_{\text{SMOW}}$ of the oceans through time.}
\label{veizer}
\end{figure}
